---
title: "In-class Exercise 12: Calibrating Geographically Weighted Predictive Models"
author: "Dr. Kam Tin Seong<br/>Assoc. Professor of Information Systems(Practice)"
institute: "School of Computing and Information Systems,<br/>Singapore Management University"
date: "`r Sys.Date()`"
format: 
  revealjs:
    width: 1600
    height: 900
    show-notes: false
    slide-number: true
    show-slide-number: all
date-format: "DD MMM YYYY"
editor: visual
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

## Getting Started

### Installing and Loading R packages

::: {style="font-size: 1.50em"}
```{r}
pacman::p_load(sf, spdep, GWmodel, SpatialML, 
               tmap, rsample, yardstick, tidyverse,
               knitr, kableExtra, spatialRF)
```
:::

## Preparing Data

:::::::: panel-tabset
### Data import

::: {style="font-size: 1.50em"}
```{r}
mdata <- read_rds("data/rds/mdata.rds")
```
:::

### Data Sampling

Calibrating predictive models are computational intensive, especially random forest method is used. For quick prototyping, a 10% sample will be selected at random from the data by using the code chunk below.

::: {style="font-size: 1.50em"}
```{r}
set.seed(1234)
HDB_sample <- mdata %>%
  sample_n(1500)
```
:::

### Checking of overlapping point

::: callout-warning
When using GWmodel to calibrate explanatory or predictive models, it is very important to ensure that there are no overlapping point features
:::

The code chunk below is used to check if there are overlapping point features.

::: {style="font-size: 1.50em"}
```{r}
overlapping_points <- HDB_sample %>%
  mutate(overlap = lengths(st_equals(., .)) > 1)
summary(overlapping_points$overlap)
```
:::

### Spatial jitter

In the code code chunk below, [`st_jitter()`](https://r-spatial.github.io/sf/reference/st_jitter.html) of **sf** package is used to move the point features by 5m to avoid overlapping point features.

::: {style="font-size: 1.50em"}
```{r}
HDB_sample <- HDB_sample %>%
  st_jitter(amount = 1)
```
:::
::::::::

## Data Sampling

The entire data are split into training and test data sets with 65% and 35% respectively by using *initial_split()* of **rsample** package. rsample is one of the package of tigymodels.

::: {style="font-size: 1.50em"}
```{r}
set.seed(1234)
resale_split <- initial_split(HDB_sample, 
                              prop = 6.67/10,)
train_data <- training(resale_split)
test_data <- testing(resale_split)
```
:::

It is always a good practice to save the training and test data sets for subsequent used. 

::: {style="font-size: 1.50em"}
```{r}
#| eval: false
write_rds(train_data, "data/rds/train_data.rds")
write_rds(test_data, "data/rds/test_data.rds")
```
:::

```{r}
#| eval: false
#| echo: false
train_data <- read_rds("data/model/train_data.rds")
test_data <- read_rds("data/model/test_data.rds")
```

------------------------------------------------------------------------

### Multicollinearity check

:::: panel-tabset
#### The code

In order to avoid multicollineariy. In the code chunk below, [`ggcorrmat()`](https://indrajeetpatil.github.io/ggstatsplot/articles/web_only/ggcorrmat.html) of **ggstatsplot** is used to plot a correlation matrix to check if there are pairs of highly correlated independent variables.

::: {style="font-size: 1.50em"}
```{r}
#| eval: false
mdata_nogeo <- mdata %>%
  st_drop_geometry()
ggstatsplot::ggcorrmat(mdata_nogeo[, 2:17])
```
:::

#### The plot

```{r}
#| echo: false
#| fig-width: 12
#| fig-height: 8
mdata_nogeo <- HDB_sample %>%
  st_drop_geometry()
ggstatsplot::ggcorrmat(mdata_nogeo[, 2:17])
```
::::

## Building a non-spatial multiple linear regression

:::: panel-tabset
### The code chunk

::: {style="font-size: 1.50em"}
```{r}
#| eval: false
price_mlr <- lm(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                data=train_data)
olsrr::ols_regress(price_mlr)
```
:::

### The report

```{r}
#| echo: false
price_mlr <- lm(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                data=train_data)
olsrr::ols_regress(price_mlr)
```
::::

------------------------------------------------------------------------

### Multicollinearity check with VIF

::::: panel-tabset
#### VIF

::: {style="font-size: 1.50em"}
```{r}
vif <- performance::check_collinearity(price_mlr)
kable(vif, 
      caption = "Variance Inflation Factor (VIF) Results") %>%
  kable_styling(font_size = 18) 
```
:::

#### Plotting VIF

::: {style="font-size: 1.50em"}
```{r}
plot(vif) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
:::
:::::

## Predictive Modelling with gwr

::::: panel-tabset
### Computing bw

::: {style="font-size: 1.50em"}
```{r}
#| output: false
gwr_bw_train_ad <- bw.gwr(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                  data=train_data,
                  approach="CV",
                  kernel="gaussian",
                  adaptive=TRUE,
                  longlat=FALSE)
gwr_bw_train_ad
```
:::

```{r}
#| eval: false
#| echo: false
write_rds(gwr_bw_train_ad, "data/model/gwr_bw_train_ad.rds")
```

```{r}
#| eval: false
#| echo: false
gwr_bw_train_ad <- read_rds("data/model/gwr_bw_train_ad.rds")
```

### Model calibration

::: {style="font-size: 1.50em"}
```{r}
#| output: false
gwr_ad <- gwr.basic(formula = resale_price ~ floor_area_sqm +
                      storey_order + remaining_lease_mths + 
                      PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                      PROX_MRT + PROX_PARK + PROX_MALL +
                      PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                      WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                      WITHIN_1KM_PRISCH,
                    data=train_data,
                    bw=20, 
                    kernel = 'gaussian', 
                    adaptive=TRUE,
                    longlat = FALSE)
```
:::

### Model output

```{r}
#| echo: false
gwr_ad
```

```{r}
#| echo: false
write_rds(gwr_ad, "data/model/gwr_ad.rds")
```

```{r}
#| echo: false
gwr_ad <- read_rds("data/model/gwr_ad.rds")
```

### Computing test data bw

::: {style="font-size: 1.50em"}
```{r}
#| eval: false
gwr_bw_test_ad <- bw.gwr(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                  data=test_data,
                  approach="CV",
                  kernel="gaussian",
                  adaptive=TRUE,
                  longlat=FALSE)
```
:::

```{r}
#| echo: false
#| eval: false
write_rds(gwr_bw_test_ad,
          "data/model/gwr_bw_test_ad.rds")
```

```{r}
#| eval: false
#| echo: false
gwr_bw_test_adaptive <- read_rds(
  "data/model/gwr_bw_test_ad.rds")
```

### Predicting with test data

::: {style="font-size: 1.50em"}
```{r}
gwr_pred <- gwr.predict(formula = resale_price ~ floor_area_sqm +
                          storey_order + remaining_lease_mths + PROX_CBD +
                          PROX_ELDERLYCARE + PROX_HAWKER + PROX_MRT + 
                          PROX_PARK + PROX_MALL + PROX_SUPERMARKET + 
                          WITHIN_350M_KINDERGARTEN + WITHIN_350M_CHILDCARE +
                          WITHIN_350M_BUS + WITHIN_1KM_PRISCH, 
                        data=train_data, 
                        predictdata = test_data, 
                        bw=20, 
                        kernel = 'gaussian', 
                        adaptive=TRUE, 
                        longlat = FALSE)

```
:::

### Saving predicted values

::: {style="font-size: 1.50em"}
```{r}
gwr_pred_df <-as.data.frame(
  gwr_pred$SDF$prediction) %>%
  rename(gwr_pred = "gwr_pred$SDF$prediction")
```
:::
:::::

## Predictive Modelling with RF method

::::::: panel-tabset
### Data preparation

Firstly, code chunk below is used to extract the coordinates of training and test data sets

::: {style="font-size: 1.50em"}
```{r}
coords <- st_coordinates(HDB_sample)
coords_train <- st_coordinates(train_data)
coords_test <- st_coordinates(test_data)
```
:::

Next, code chunk below is used to drop the geometry column of both training and test data sets.

::: {style="font-size: 1.50em"}
```{r}
train_data_nogeom <- train_data %>%
  st_drop_geometry()
```
:::

### Calibrating RF model

::: {style="font-size: 1.50em"}
```{r}
#| output: false
set.seed(1234)
rf <- ranger(resale_price ~ floor_area_sqm + storey_order + 
               remaining_lease_mths + PROX_CBD + PROX_ELDERLYCARE + 
               PROX_HAWKER + PROX_MRT + PROX_PARK + PROX_MALL + 
               PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
               WITHIN_350M_CHILDCARE + WITHIN_350M_BUS + 
               WITHIN_1KM_PRISCH,
             data=train_data_nogeom)

```
:::

### Model output

::: {style="font-size: 1.50em"}
```{r}
rf
```
:::

### Preparing the test data

::: {style="font-size: 1.50em"}
```{r}
test_data_nogeom <- cbind(
  test_data, coords_test) %>%
  st_drop_geometry()
```
:::

### Predicting with rf

::: {style="font-size: 1.50em"}
```{r}
rf_pred <- predict(rf, 
                   data = test_data_nogeom)
```
:::

### Saving the predicted values

The output of `predict()` of **ranger** is object of class *ranger* with elements.  Code chunk below is used to extract values from *predictions* of predicted  classes and save the values in a data frame called *rf_pred_df*. 

::: {style="font-size: 1.50em"}
```{r}
rf_pred_df <- as.data.frame(rf_pred$predictions) %>%
  rename(rf_pred = "rf_pred$predictions")
```
:::

:::::::

## Predictive Modelling with SpatialML

::::: panel-tabset
### Determining bandwidth

::: {style="font-size: 1.50em"}
```{r}
#| eval: false
set.seed(1234)
grd_bw_train_ad <- grf.bw(formula = resale_price ~ floor_area_sqm + storey_order +
                            remaining_lease_mths + PROX_CBD + PROX_ELDERLYCARE +
                            PROX_HAWKER + PROX_MRT + PROX_PARK + PROX_MALL +
                            PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                            WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                            WITHIN_1KM_PRISCH,
                          dataset=train_data_nogeom, 
                          kernel="adaptive",
                          coords=coords_train)
```
:::

### Calibrating with grf

::: {style="font-size: 1.50em"}
```{r}
#| output: false
set.seed(1234)
grf_ad <- grf(formula = resale_price ~ floor_area_sqm + storey_order +
                remaining_lease_mths + PROX_CBD + PROX_ELDERLYCARE + 
                PROX_HAWKER + PROX_MRT + PROX_PARK + PROX_MALL + 
                PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                WITHIN_1KM_PRISCH,
              dframe=train_data_nogeom, 
              bw=20,
              kernel="adaptive",
              coords=coords_train)
```
:::

### Model output

::: {style="font-size: 1.50em"}
```{r}
grf_ad
```
:::

```{r}
#| echo: false
#| eval: false
write_rds(grf_ad, "data/model/grf_ad.rds")
```

```{r}
#| echo: false
#| eval: false
grf_ad <- read_rds("data/model/grf_ad.rds")
```

### Preparing the test data

::: {style="font-size: 1.50em"}
```{r}
test_data_nogeom <- cbind(
  test_data, coords_test) %>%
  st_drop_geometry()
```
:::

### Predicting with the test data

In the code chunk below, `predict.grf()` of **spatialML** for predicting re-sale prices in the **test data** set (i.e. *test_data_nogeom*)

::: {style="font-size: 1.50em"}
```{r}
grf_pred <- predict.grf(grf_ad, 
                         test_data_nogeom,
                         x.var.name="X",
                         y.var.name="Y", 
                         local.w=1,
                         global.w=0)
```
:::

### Saving the predicted values 

Next, the code chunk below is used to convert the output from *grf_pred* into a data.frame.

::: {style="font-size: 1.50em"}
```{r}
grf_pred_df <- as.data.frame(grf_pred)
```
:::
::::::::

## Model Comparison

::: panel-tabset
### Combining models' outputs

::: {style="font-size: 1.50em"}
```{r}
test_data_pred <- test_data %>%
  select(resale_price) %>%
  cbind(gwr_pred_df) %>%
  cbind(rf_pred_df) %>%
  cbind(grf_pred_df)
```
:::

### Transposing data 

::: {style="font-size: 1.50em"}
```{r}
test_longer <- test_data_pred %>%
  st_drop_geometry() %>%
  pivot_longer(cols = ends_with("pred"), 
               names_to = "model", 
               values_to = "predicted")
```
:::

### Renaming

::: {style="font-size: 1.50em"}
```{r}
model_labels <- c(
  gwr_pred = "gwr",
  rf_pred = "Random Forest",
  grf_pred = "gwRF")

test_longer <- test_longer %>%
  mutate(model = recode(
    model, !!!model_labels))
```
:::

### Computing rmse

::: {style="font-size: 1.50em"}
```{r}
rmse_results <- test_longer %>%
  group_by(model) %>%
  rmse(truth = resale_price, 
       estimate = predicted) %>%
  rename(rmse = .estimate) %>%
  select(model, rmse)
```
:::
:::

## Model Comparison Plots

::: panel-tabset

### Bar chart

::: columns
::: {.column width=50%}
::: {style="font-size: 1.50em"}
```{r}
#| eval: false
ggplot(rmse_results, 
       aes(x = reorder(model, rmse), 
           y = rmse, 
           fill = "skyblue")) +
  geom_bar(stat = "identity", 
           fill = "skyblue", 
           color = "black", 
           width = 0.7) +
  labs(title = "RMSE Comparison of Models", 
       y = "RMSE", 
       x = "Model") +
  theme_minimal()
```
:::
:::

::: {.column width=50%}
```{r}
#| echo: false
ggplot(rmse_results, 
       aes(x = reorder(model, rmse), 
           y = rmse, 
           fill = "skyblue")) +
  geom_bar(stat = "identity", 
           fill = "skyblue", 
           color = "black", 
           width = 0.7) +
  labs(title = "RMSE Comparison of Models", 
       y = "RMSE", 
       x = "Model") +
  theme_minimal()
```
:::
:::
:::

---

### Scatter plots

:::: panel-tabset

#### The code chunk

::: {style="font-size: 1.25em"}
```{r}
#| eval: false
test_longer <- test_longer %>%
  left_join(rmse_results, 
            by = "model")

ggplot(data = test_longer,
       aes(x = predicted,
           y = resale_price)) +
  facet_wrap(~ model) +
  geom_point() +
  geom_text(data = test_longer, 
            aes(x = Inf, y = Inf, 
                label = paste("RMSE:", round(rmse, 2))),
            hjust = 1.1, vjust = 1.1, 
            color = "black", size = 4)
```
:::

#### The plot

```{r}
#| echo: false
#| fig-width: 14
test_longer <- test_longer %>%
  left_join(rmse_results, 
            by = "model")

ggplot(data = test_longer,
       aes(x = predicted,
           y = resale_price)) +
  facet_wrap(~ model) +
  geom_point() +
  geom_text(data = test_longer, 
            aes(x = Inf, y = Inf, 
                label = paste("RMSE:", round(rmse, 2))),
            hjust = 1.1, vjust = 1.1, 
            color = "black", size = 4)
```

::::

## Variable Importance

::: panel-tabset

### Preparing the output

::: {style="font-size: 1.50em"}
```{r}
var_imp <- data.frame(
  Variable = names(grf_ad$Global.Model$variable.importance),
  Importance = grf_ad$Global.Model$variable.importance
)
```
:::

### The plot

```{r}
#| echo: false
#| fig-height: 7
ggplot(var_imp, aes(x = reorder(Variable, 
                                Importance), 
                    y = Importance)) +
  geom_bar(stat = "identity", 
           fill = "skyblue") +
  coord_flip() +
  labs(
    title = "Variable Importance from Ranger Model",
    x = "Variables",
    y = "Importance"
  ) +
  theme_minimal()
```

### The code

::: {style="font-size: 1.50em"}
```{r}
#| eval: false
ggplot(var_imp, aes(x = reorder(Variable, 
                                Importance), 
                    y = Importance)) +
  geom_bar(stat = "identity", 
           fill = "skyblue") +
  coord_flip() +
  labs(
    title = "Variable Importance from Ranger Model",
    x = "Variables",
    y = "Importance"
  ) +
  theme_minimal()
```
:::
:::